Research Proposal

### 2012-03-13 dissertation-proposal/research-proposal.tex
Start on second draft

\documentclass[12pt, a4paper]{article}
\setlength{\oddsidemargin}{0.5cm}
\setlength{\evensidemargin}{0.5cm}
\setlength{\topmargin}{-1.6cm}
\setlength{\leftmargin}{0.5cm}
\setlength{\rightmargin}{0.5cm}
\setlength{\textheight}{24.00cm} 
\setlength{\textwidth}{15.00cm}
\parindent 0pt
\parskip 5pt
\pagestyle{plain}

@@

Programmers have long been divided between statically typed and 
dynamically typed languages.
Traditionally, languages have been defined as using one of these styles,
but recently languages such as Racket/Typed Racket, Dart and Qi/Shen
have borrowed from both.

@@

There has been little practical experience writing programs in a single
language that combine both styles.  One example is
Typed Racket, which provides a statically typed variation of Racket 
with the intention of safely porting untyped code to typed code. 
Typed code can interact with untyped code while preserving some properties
of typed code by, for example, delaying static type checks until runtime.
Porting untyped Racket code to Typed Racket generally involves 
adding type annotations to top level declarations with minimal (usually zero)
modification to existing code structure.

@@

I hypothesise that adding an optional static type system to a dynamically typed language
will increase its utility, while avoiding some disadvantages of statically typed
languages.

I expect to develop a library for Clojure which provides an optional static
type system, intended for practical use.

The rich type annotations that will be accumulated could also be
useful elsewhere, such as a function search engine like
(name of haskell search engine? TODO)

@@

>>>>

\documentclass[12pt, a4paper]{article}

\usepackage{biblatex}
\bibliography{bibliography}

\setlength{\oddsidemargin}{0.5cm}
\setlength{\evensidemargin}{0.5cm}
\setlength{\topmargin}{-1.6cm}
\setlength{\leftmargin}{0.5cm}
\setlength{\rightmargin}{0.5cm}
\setlength{\textheight}{24.00cm} 
\setlength{\textwidth}{15.00cm}
\parindent 0pt
\parskip 5pt
\pagestyle{plain}

@@

Dynamic languages are designed to be convenient for the programmer to
write programs quickly and aspire to get out of the programmer's
way as much as possible. When programs grow large and stabilize,
the features of static languages are missed, specifically static
type checking.

Having a dynamic language that supports static properties 
has many potential practical benefits. 
Module-by-module porting of existing untyped code to a typed language 
is the largest benefit cited by Tobin-Hochstadt \cite{TypedScheme:2010}.
He developed Typed Scheme
\cite{TypedScheme:2010} 
to safely and incrementally port existing Scheme code, a dynamic language, to
Typed Scheme, a static language.

A recent example is Dart 
\cite{Dart:2012:Online}, a dynamically typed language that allows optional
static checks.

Good type systems deeply integrate with the language they are written for.
Statically typed languages usually include a type system in early releases,
which is able to evolve with the language while it is malleable.

The type system understands common idioms
and becomes the programmer's ally.

@@

Typed Racket \cite{TypedScheme:2010}
, which provides a statically typed variation of Racket 
with the intention of safely porting untyped code to typed code. 
Typed code can interact with untyped code while preserving some properties
of typed code by, for example, delaying static type checks until runtime.
Porting untyped Racket code to Typed Racket generally involves 
adding type annotations to top level declarations with minimal (usually zero)
modification to existing code structure.

@@

My goal is to develop an optional static type system for Clojure, 
intended for practical use, to be provided as a library.

It will be based on the lessons learnt throught the development
of Typed Scheme, and as a response to Tobin-Hochstadt's \cite{TypedScheme:2010}
suggestion to add type systems to dynamic languages.

I expect this library to be an important asset to Clojure's feature set.
I expect it will become a factor in language decision and justification for programming projects.

@@

\printbibliography
